---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include = FALSE
 
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pcapi

<!-- badges: start -->
[![R-CMD-check](https://github.com/RaoulWolf/pcapi/workflows/R-CMD-check/badge.svg)](https://github.com/RaoulWolf/pcapi/actions)
<!-- badges: end -->


The goal of the pcapi package is to provide a minimal and lightweight 
interface to the [PubChem](https://pubchem.ncbi.nlm.nih.gov/) API services. 

The dependencies of pcapi are kept at a bare minimum: 
[curl](https://cran.r-project.org/web/packages/curl/index.html) for handling 
the API requests and 
[jsonlite](https://cran.r-project.org/web/packages/jsonlite/index.html) to 
parse data from JSON format.

## Installation

You can install the development version of pcapi from 
[GitHub](https://github.com/) with:

```{r install}
#| eval = FALSE

# install.packages("devtools")
remotes::install_github("RaoulWolf/pcapi")
```

## Examples

The following examples show the functionality of two API functions of
pcapi: `post_to_cid()` and `post_to_property()`.

This is an example which shows you how to get the PubChem CID for aspirin:

```{r example1}
#| eval = FALSE

library(pcapi)
x <- "aspirin"
post_to_cid(x, format = "name")
```

Here is an example which shows you how to get the properties for a PubChem CID. 
In this case, again, for aspirin:

```{r example2}
#| eval = FALSE

cid <- 2244
post_to_properties(cid)
```

## Acknowledgement

This R package was developed by the EnviData initiative at the 
[Norwegian Geotechnical Institute (NGI)](https://www.ngi.no/eng) as part of the 
project 
[ZeroPM: Zero pollution of Persistent, Mobile substances](https://zeropm.eu/). 
This project has received funding from the European Union's Horizon 2020 
research and innovation programme under grant agreement No 101036756.

---

If you find this package useful and can afford it, please consider making a 
donation to a humanitarian non-profit organization, such as 
[Sea-Watch](https://sea-watch.org/en/). Thank you. 
