---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include = FALSE
 
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pcapi <img src="man/figures/logo.svg" align="right" height="139" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/RaoulWolf/pcapi/workflows/R-CMD-check/badge.svg)](https://github.com/RaoulWolf/pcapi/actions)
[![Codecov test coverage](https://codecov.io/gh/RaoulWolf/pcapi/branch/master/graph/badge.svg)](https://app.codecov.io/gh/RaoulWolf/pcapi?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

> A [ZeroPM](https://zeropm.eu/) R package

The goal of the pcapi package is to provide a minimal and lightweight 
interface to the [PubChem](https://pubchem.ncbi.nlm.nih.gov/) API services. 

The dependencies of pcapi are kept at a bare minimum: 
[curl](https://cran.r-project.org/web/packages/curl/index.html) for handling 
the API requests and 
[jsonlite](https://cran.r-project.org/web/packages/jsonlite/index.html) to 
parse data from JSON format.

Currently, the following formats can be used to retrieve PubChem 
CIDs: InChI, InChIKey, names (including synonyms), SDF, and SMILES. 

### But what about [webchem](https://docs.ropensci.org/webchem/)?

First of all, [webchem](https://docs.ropensci.org/webchem/) is a fantastic 
package - please use it! webchem offers support for many different API 
services, not only PubChem's API service. However, there are issues related to 
the way webchem sets up the queries against PubChem API services, which can 
result in errors. This has been observed, for example, when SMILES or InChI 
strings contain backslashes, or an MOL/SDF is used as input. With pcapi, these 
problems are (hopefully) solved. Additionally, pcapi will continuously 
implement some of the lesser known functionalities, such as the retrival of 
transformation products. 

## Installation

You can install the development version of pcapi from 
[GitHub](https://github.com/) with:

```{r install}
#| eval = FALSE

# install.packages("devtools")
remotes::install_github("RaoulWolf/pcapi")
```

## Functions

PubChem's API services are vast and flexible. Not all functionality will be 
available immediately, and some functionalities may not be worth it to 
implement. That being said, the goal of pcapi is to provide a satisfying 
user-experience, covering the most relevant functionalities of the API. The 
following table gives an overview of the available functions, as well as their 
expected inputs and outputs. 

| Function                   | Expected input                                   | Output                                |
|:-------------------------- |:------------------------------------------------ |:------------------------------------- |
| `post_to_cid()`            | InChI, InChIKey, MOL, name, SDF, SMILES, synonym | integer vector of PubChem CID(s)      |
| `post_to_property()`       | PubChem CID                                      | data frame of properties              |
| `post_to_transformation()` | PubChem CID                                      | data frame of transformation products |
| `post_to_use()`            | PubChem CID                                      | data frame of use categories          |

## Examples

The following examples show the functionality of two API functions of
pcapi: `post_to_cid()` and `post_to_property()`.

This is an example which shows you how to get the PubChem CID for the herbicide 
[atrazine](https://en.wikipedia.org/wiki/Atrazine):

```{r example1}
library(pcapi)

atrazine_cid <- post_to_cid("atrazine", format = "name")
atrazine_cid
```

Using this PubChem CID, we can now get all available descriptors and 
physico-chemical properties for atrazine: 

```{r example2}
atrazine_properties <- post_to_property(atrazine_cid, property = "all")
str(atrazine_properties)
```

To get an overview of all transformation products, the PubChem CID can be used 
as well:

```{r example3}
atrazine_transformations <- post_to_transformation(atrazine_cid)
str(atrazine_transformations)
```

The use categories can also be retrieved using the PubChem CID:

```{r example4}
atrazine_uses <- post_to_use(atrazine_cid)
str(atrazine_uses)
```


## Acknowledgement

This R package was developed by the EnviData initiative at the 
[Norwegian Geotechnical Institute (NGI)](https://www.ngi.no/eng) as part of the 
project 
[ZeroPM: Zero pollution of Persistent, Mobile substances](https://zeropm.eu/). 
This project has received funding from the European Union's Horizon 2020 
research and innovation programme under grant agreement No 101036756.

---

If you find this package useful and can afford it, please consider making a 
donation to a humanitarian non-profit organization, such as 
[Sea-Watch](https://sea-watch.org/en/). Thank you. 
